Effectful Semantics for Substructural SSA

### 2023-11-17 pldi24/paper.tex
Tap

We parametrize our syntax by a set of \textit{base types} \(X \in \mc{T}\) and,
for each pair of types \(A, B\), a set of \textit{instructions} \(f \in
\mc{I}^q_p\) with \textit{centrality} \(p \in \{0, 1\}\) and \textit{quantity}
\(q \subseteq \tint\). As described above, a \textit{central} instruction is one
for which execution order does not matter, while the quantity \(q\) indicates
whether it is safe to delete an instruction if its result is unused (in which
case the instruction is \textit{affine}, and \(\taff \subseteq q\)) or duplicate
an instruction used multiple times (in which case the instruction is
\textit{relevant}, and \(\trel \subseteq q\)). An instruction which is both
affine and relevant is called \textit{pure}. The set of types is generated by
base types \(X\), the \textit{unit type} \(\mb{1}\), the \textit{boolean type}
\(\mb{2}\), and tensor products of types \(A \otimes B\). We write \(\otimes\)
to underscore the fact that types are allowed to be substructural: similarly to
instructions, every type is assigned a \textit{linearity} \(\ms{lin}(A)
\subseteq \tint\) satisfying the constraints in Figure \ref{fig:ssa-structural},
except that a type which is both affine and relevant is called
\textit{intuitionistic}. In particular, the rules \rle{unit-lin} and
\rle{bool-lin} state that both the unit and boolean types must be
intuitionistic; the rule \rle{pair-lin} says that the linearity of a pair is
simply the intersection of the linearity of its components. 
% The rules \rle{nil-lin} and \rle{cons-lin} together say that the
% linearity of a context is the intersection of the linearity of its variables,
% where the linearity of a variable is the intersection of the linearity of its
% type and its quantity annotation.

Our syntax is composed of three main syntactic categories:
\begin{itemize}
  \item \textit{Terms} \(a, b, c, d, e\), which are typed via the judgement
  \(\hasty{\Gamma}{p}{a}{A}{q}\), which says that the term \(a\) is of type
  \(A\) with \textit{centrality} \(p = 1\) if \textit{central}, or \(p = 0\)
  otherwise, and \textit{quantity} \(q \subseteq \{\ms{a}, \ms{r}\}\). So the
  centrality \(p\) indicates whether it is safe to substitute this term for a
  variable, and \(q\), plus the linearity of \(A\), control how many times that
  variable can occur: if \(\taff \subseteq q\) then the variable is called
  \textit{affine} or \textit{discardable} and it is permissible to leave it
  unused, whereas if \(\trel \subseteq q\) it is called \textit{relevant} or
  \textit{duplicable} and it is permissible to use multiple times. A term which
  is central, relevant, and affine is called \textit{pure}. The rules for this
  judgement are given in Figure \ref{fig:ssa-term-typing}.

@@

Here \textit{contexts} \(\Gamma\), consist of a list of variables \(x\) given a
type \(A\) annotated with a \textit{quantity} \(q \subseteq \tint\). A variable
\(x: A^q\) is called \textit{affine} if its type \(A\) is affine (i.e.
\(\tlin{A}{\taff}\)) \textit{and} \(\taff \subseteq q\), in which case it can be
freely ignored. Similarly, \(x\) is \textit{relevant} if its type \(A\) relevant
(i.e. \(\tlin{A}{\trel}\)) \textit{and} \(\trel \subseteq q\), in which case it
can be freely ignored. This is formalized in the structural judgement
\(\csplits{\Gamma}{\Delta}{\Xi}\), which says that \(\Gamma\) can be split into
subcontexts \(\Delta\) and \(\Xi\) (weakening is simply the special case
\(\cwk{\Gamma}{\Delta} \iff \csplits{\Gamma}{\Delta}{\cdot}\)), as per the
typing rules in Figure \ref{fig:ssa-structural}. In particular, the rule
\rle{split-nil} provides a base case by saying that the empty context can be
split into two empty contexts, while the rules \rle{split-left} and
\rle{split-right} send \(x: A^q\) to the left or right subcontext repsectively;
both these rules also allow us to reduce the quantity \(q\) to a quantity \(r
\subseteq q\), "forgetting" some of our capabilities. The rule \rle{split-dup}
allows the duplication of a variable which has a relevant type \(A\)
\textit{and} a relevant quantity \(q\), with the option to weaken the quantity
to \textit{different} quantities \(r\) and \(s\) on the left-hand side and
right-hand sides respectively. Finally, the rule \rle{split-aff} simply allows
us to ignore a variable \(x: A^q\) if \(A\) and \(q\) are affine.

\textit{label-sets} \(\ms{L}\) are defined to consist of a list of
\textit{labels} \(\lbl{\ell}\) annotated with a context \(\Gamma\) and an
argument type \(A\). The judgement \(\lwk{\ms{L}}{\ms{K}}\) states that
\(\ms{L}\) is a subcontext of \(\ms{K}\), i.e., that every program which targets
\(\ms{L}\) (i.e., finishes by either nontermination or branching to one of the
labels \(\llhyp{\ell}{\Gamma}{A} \in \ms{L}\) with live variables \(\Gamma\) and
an argument of type \(A\)) can also be viewed as targeting \(\ms{K}\). The
conditions for this are formalized in Figure \ref{fig:ssa-structural} as
follows: the rule \rle{join-nil} provides a base case stating that the empty
label weakens into itself. The rule \rle{join-cons} says that if \(\ms{L}\)
weakens into \(\ms{K}\) and \(\Delta\) weakens into \(\Gamma\), then
\(\lwk{\ms{L}, \llhyp{\ell}{\Gamma}{A}}{\ms{K}, \llhyp{\ell}{\Delta}{A}}\),
allowing us to "forget" any extra variables dropped by \(\cwk{\Gamma}{\Delta}\).
Finally, the rule \rle{join-zero} simply says that if \(\lwk{\ms{L}}{\ms{K}}\),
then \(\ms{L}\) also weakens to any extension of \(\ms{K}\). Note that this
judgement is "backwards," in that the "larger" context is on the \textit{right};
this corresponds to the placement of the label context on the right-hand side of
the typing judgement for blocks, and is by analogy to our semantics for blocks
and CFGs in Section \ref{sec:semantics}.

@@

We can now get to the typing judgement for expressions in Figure
\ref{fig:ssa-term-typing}. The rule \rle{var} states that \(x: A^q\) is
well-typed in \(\Gamma\) if all other variables are affine (i.e., can be
dropped) and \(q \subseteq q'\), where \(x: A^{q'} \in \Gamma\); this reflects
that we're allowed to \textit{forget} capabilities, but not \textit{add} them.
The \rle{app} rule states that, given an instruction \(f\) which takes in an
argument \(A\) and has a return type of \(B\) and a \textit{central} term
\(\hasty{\Gamma}{1}{a}{A}{q}\), we can conclude that
\(\hasty{\Gamma}{p}{f\;a}{B}{q}\). We, of course, require that \(f\) itself have
centrality \(p\) and quantity \(q\), i.e. that \(f \in \mc{I}_p^q(A, B)\). The
\rle{pair} rule says that we can type a pair \((a, b)\) in \(\Gamma\) if
\(\csplits{\Gamma}{\Delta}{\Xi}\) where \(\Delta\) types \(a\) and \(\Xi\) types
\(b\); the quantity of a pair is simply the intersection of the quantity of its
components (this is expressed in the rule by allowing any quantity which types
\textit{both} components), and both components must be central. Finally, we have
the rules \rle{let} and \rle{let2}, correspond to let-bindings and destructuring
let-bindings respectively. In both cases, we split the context \(\Gamma\) into
subcontext \(\Delta, \Xi\), one of which is used to type a central subterm
\(a\), which is bound to variables \(x\) or destructured to variables \(x, y\)
respectively, in a potentially non-central term \(e\) which is otherwise typed
in \(\Delta\).

@@

The rules for block typing are given in Figure \ref{fig:ssa-block-typing}.
\rle{br} says that, given a pure term \(a: A\) (that is, central, relevant, and
affine), we can type a block \(\lbrb{\ell}{a}\) which branches to \(\ell\) with
argument \(a\), where \(\ms{L}\) contains the context \(\ell\) with an argument
type of \(A\) and an appropriate set of unused variables. The rule \rle{ite}
typechecks if-then-else statements by stating that, if \(\Gamma\) splits into
\(\Delta\) and \(\Xi\), where \(\Xi\) typechecks a pure, intuitionistic
discriminant \(e: \mb{2}\) and \(\Delta\) typechecks both \(s, t\) to branch
into \(\ms{L}\), then the whole if-statement \(\ite{e}{s}{t}\) also typechecks
as branching into \(\ms{L}\) in constext \(\Gamma\). \rle{let-blk} and
\rle{let2-blk} are quite similar to their term analogues, except that the
bound term \(e\) is allowed to be impure (otherwise, programs could only
make use of pure functions!), and the remainder of the context plus the bound
variable is used to type a block rather than a term. Finally, probably the most
interesting rule is \rle{where}, which says that, given a block \(t\) which
branches to \(\ms{L}\) in context \(\Gamma\), and labels \(L\) mapping
\(\ms{L}\) to \(\ms{K}\), \(\ewhere{t}{L}\) branches to \(\ms{K}\). Here, \(L\)
mapping \(\ms{L}\) to \(\ms{K}\) is defined using the typing rules \rle{nil-br},
which says that an empty control-flow graph simply forwards the input labels
directly to the output labels, and \rle{cons-br}, which says that if \(L\) maps
\(\ms{L}\) to either \(\ms{K}\) or the label \(\lbl{\ell}[\Gamma](A)\), and
\(t\) branches to \(\ms{L}\) in the context \(\Gamma, \thyp{x}{A}{}\), then
\(L\) maps \(\ms{L}\) to \(\ms{K}\), having (potentially recursively) defined
\(\lbl{\ell}\).

@@

To solve this problem, we need to introduce the concept of a
\textit{label-substitution} \(\mc{L}\), which we define as a function that takes
labels \(\lbl{\ell}\) and variables \(x\) to blocks \(t\) such that, for any
fresh variable \(x\), \([a/x]\mc{L}(\lbl{\ell}, x) = \mc{L}(\lbl{\ell}, a)\). We
introduce the judgement \(\lbsubst{\mc{L}}{\ms{L}}{\ms{K}}\) to state that the
label-substitution \(\mc{L}\) takes the label-set \(\ms{L}\) to the label-set
\(\ms{K}\); that is, for each \(\lbl{\ell}[\Gamma](x: A) \in \ms{L}\), we have
that \(\mc{L}(\lbl{\ell}, x)\) is a block that targets \(\ms{K}\). This is
formally defined via the rules \rle{lb-subst-nil}, which states that this is
vacuously true for \(\ms{L} = \cdot\), and \rle{lb-subst-cons}, which says that
if \(\lbsubst{\mc{L}}{\ms{L}}{\ms{K}}\) and \(\haslb{\Gamma,
\thyp{x}{A}{}}{\mc{L}(\lbl{\ell}, x)}{\ms{K}}\) then \(\lbsubst{\mc{L}}{\ms{L},
\lhyp{\lbl{\ell}}{\Gamma}{A}}{\ms{K}}\), in Figure \ref{fig:ssa-subst-typing}.
Given a context \(\gamma\) and a label-set \(\ms{L}\), we can convert it to a
label-substitution \(\substlbs{\gamma}{\ms{L}}\) using the definition in Figure
\ref{fig:ssa-subst-ops}: using this, we can state the full substitution lemma as
follows:
\begin{lemma}[Substitution] 
  Given \(\issubst{\gamma}{\Theta}{\Gamma}\),
  \begin{itemize}
    \item If \(\hasty{\Gamma}{p}{a}{A}{q}\), then
    \(\hasty{\Theta}{p}{[\gamma]a}{A}{q}\) and
    \(\hasty{\Theta}{p}{\exprletsubst{\substctx{\gamma}{\Gamma}}{a}}{A}{q}\)
    \item If \(\haslb{\Gamma}{t}{\ms{L}}\), then there exists \(\ms{K}\) such
    that \(\lbsubst{\substlbs{\gamma}{\ms{L}}}{\ms{K}}{\ms{L}}\) and
    \(\haslb{\Theta}{[\gamma]t}{\ms{K}}\),
    \(\haslb{\Theta}{\stmtletsubst{\substctx{\gamma}{\Gamma}}{t}}{\ms{K}}\)
    \item If \(\lhaslb{\ms{W}}{L}{\ms{L}}\), then there exists \(\ms{K},
    \ms{W}'\) such that 
    \begin{itemize}
      \item \(\lbsubst{\substlbs{\gamma}{\ms{L}}}{\ms{K}}{\ms{L}}\) 
      \item \(\lbsubst{\substlbs{\gamma}{\ms{W}'}}{\ms{W}}{\ms{W}'}\) 
      \item \(\lhaslb{\ms{W}'}{[\gamma]L}{\ms{K}}\)
    \end{itemize}
  \end{itemize}
  Similarly, given \(\lbsubst{\mc{L}}{\ms{L}}{\ms{K}}\), we have that for all
  \(\haslb{\Gamma}{t}{\ms{L}}\), \(\haslb{\Gamma}{[\mc{L}]t}{\ms{K}}\)
\end{lemma}

>>>>

We parametrize our syntax by a set of \textit{base types} \(X \in \mc{T}\) and,
for each pair of types \(A, B\), a set of \textit{instructions} \(f \in
\mc{I}^q_p(A, B)\) with \textit{centrality} \(p = 1\) if \textit{central} or \(p
= 0\) otherwise and \textit{quantity} \(q \subseteq \tint\). As discussed above,
an instruction is considered \textit{central} if execution order does not
matter, while the quantity \(q\) indicates whether it is safe to delete an
instruction if its result is unused (in which case the instruction is
\textit{affine}, and \(\taff \subseteq q\)) or duplicate an instruction used
multiple times (in which case the instruction is \textit{relevant}, and \(\trel
\subseteq q\)). An instruction which is, central, relevant, and affine is called
\textit{pure}. Note that we assume that \(p \leq p' \and q \subseteq r \implies
\mc{I}_p^q(A, B) \supseteq \mc{I}_{p'}^{r}(A, B)\). The set of types is
generated by base types \(X\), the \textit{unit type} \(\mb{1}\), the
\textit{boolean type} \(\mb{2}\), and tensor products of types \(A \otimes B\).
We write \(\otimes\) (rather than \(\times\)) to underscore the fact that types
are allowed to be substructural: given a linearity \(\ms{lin}(X) \subseteq
\tint\), we can assign every type a \textit{linearity} \(\ms{lin}(A) \subseteq
\tint\) by taking \(\ms{lin}(A \otimes B) = \ms{lin}(A) \cap \ms{lin}(B)\) and
\(\ms{lin}(\mb{1}) = \ms{lin}(\mb{2}) = \tint\). We say a type with \(\taff
\subseteq \ms{lin}(A)\) is \textit{affine}, while a type with \(\trel \subseteq
\ms{lin}(A)\) is \textit{relevant}; a type which is both affine and relevant is
called \textit{intuitionistic}. Our syntax is composed of three main syntactic
categories:
\begin{itemize}
  \item \textit{Terms} \(a, b, c, d, e\), which are typed via the judgement
  \(\hasty{\Gamma}{p}{a}{A}{q}\), which says that the term \(a\) is of type
  \(A\) with \textit{centrality} \(p\) and \textit{quantity} \(q \subseteq
  \{\ms{a}, \ms{r}\}\). Here, the centrality \(p\) indicates whether it is safe
  to substitute this term for a variable, and \(q\), plus the linearity of
  \(A\), control how many times that variable can occur: if \(\taff \subseteq
  q\) then the variable is called \textit{affine} or \textit{discardable} and it
  is permissible to leave it unused, whereas if \(\trel \subseteq q\) it is
  called \textit{relevant} or \textit{duplicable} and it is permissible to use
  multiple times. As for instructions, a term which is central, relevant, and
  affine is called \textit{pure}. The rules for this judgement are given in
  Figure \ref{fig:ssa-term-typing}.

@@

Here \textit{contexts} \(\Gamma\), consist of a list of variables \(x\) given a
type \(A\) annotated with a \textit{quantity} \(q \subseteq \tint\) (if \(q\) is
omitted, we assume \(q = \tint\)). The \textit{linearity} of a variable
\(\ms{lin}(\thyp{x}{A}{q})\) is defined to be the intersection of the linearity
of its type \(\ms{lin}(A)\) and its quantity \(q\). We say a variable \(x: A^q\)
is \textit{affine} if \(\tlin{\thyp{x}{A}{q}}{\taff}\), in which case it can be
freely ignored. Similarly, \(x\) is \textit{relevant} if
\(\tlin{\thyp{x}{A}{q}}{\trel}\), in which case it can be used multiple times.
This is formalized in the structural judgement
\(\csplits{\Gamma}{\Delta}{\Xi}\), which says that \(\Gamma\) can be split into
subcontexts \(\Delta\) and \(\Xi\) (weakening is simply the special case
\(\cwk{\Gamma}{\Delta} \iff \csplits{\Gamma}{\Delta}{\cdot}\)), as per the
typing rules in Figure \ref{fig:ssa-structural}. In particular, the rule
\rle{split-nil} states that the empty context can be split into two empty
contexts, while the rules \rle{split-left} and \rle{split-right} send \(x: A^q\)
to the left or right subcontext repsectively; both these rules also allow us to
reduce the quantity \(q\) to a quantity \(r \subseteq q\), "forgetting" some of
our capabilities. The rule \rle{split-dup} allows the duplication of a relevant
variable, with the option to weaken the quantity to \textit{different}
quantities \(r\) and \(s\) on the left and right respectively. Finally, the rule
\rle{split-aff} simply allows us to drop an affine variable.

On the other hand, \textit{label-sets} \(\ms{L}\) are defined to consist of a
list of \textit{labels} \(\lbl{\ell}\) annotated with a context \(\Gamma\)
(consisting of variables which must be live on entry to \(\lbl{\ell}\)) and an
argument type \(A\). The judgement \(\lwk{\ms{L}}{\ms{K}}\) states that
\(\ms{L}\) is a subcontext of \(\ms{K}\), i.e., has more labels with fewer live
variables, and hence that every program which targets \(\ms{L}\) can also be
viewed as targeting \(\ms{K}\). The conditions for this are formalized in Figure
\ref{fig:ssa-structural} as follows: the rule \rle{join-nil} states that the
empty label-set is a subcontext of itself. The rule \rle{join-cons} says that if
\(\ms{L}\) is a subcontext of \(\ms{K}\) and \(\Delta\) weakens into \(\Gamma\),
then \(\lwk{\ms{L}, \llhyp{\ell}{\Gamma}{A}}{\ms{K}, \llhyp{\ell}{\Delta}{A}}\),
allowing us to "forget" any extra (affine) variables dropped by
\(\cwk{\Gamma}{\Delta}\). Finally, the rule \rle{join-zero} simply says that if
\(\lwk{\ms{L}}{\ms{K}}\), then \(\ms{L}\) is also a subcontext of any extension
of \(\ms{K}\). 
% Note that this judgement is "backwards," in that the "larger" context is on
% the \textit{right}; this corresponds to the placement of the label context on
% the right-hand side of the typing judgement for blocks, and is by analogy to
% our semantics for blocks and CFGs in Section \ref{sec:semantics}.

@@

We can now get to the typing judgement for expressions in Figure
\ref{fig:ssa-term-typing}. \rle{var} states that \(\thyp{x}{A}{q}\) is
well-typed in \(\Gamma\) if all other variables are affine and \(q \subseteq
q'\), where \(\thyp{x}{A}{q'} \in \Gamma\); allowing us to \textit{forget}
capabilities, but not \textit{add} them. rle{app} states that, given an
instruction \(f \in \mc{I}_p^q(A, B)\) and a \textit{central} term
\(\hasty{\Gamma}{1}{a}{A}{q}\), we can type \(\hasty{\Gamma}{p}{f\;a}{B}{q}\).
The \rle{pair} rule says that we can type a pair \((a, b)\) in \(\Gamma\) if
\(\csplits{\Gamma}{\Delta}{\Xi}\) where \(\Delta\) types \(a\) and \(\Xi\) types
\(b\); the quantity of a pair is simply the intersection of the quantity of its
components (expressed by allowing any quantity which types \textit{both}
components), which must both be central. Finally, \rle{let} and \rle{let2},
allow us to type (destructuring) let-bindings. In both cases, we split the
context \(\Gamma\) into subcontext \(\Delta, \Xi\), one of which is used to type
a central term \(a\), which is bound to fresh variables appearing in a
potentially non-central term \(e\) which may also depend on variables in
\(\Delta\).

@@

The rules for block typing are given in Figure \ref{fig:ssa-block-typing}.
\rle{br} says that if \(\lwk{\lhyp{\ell}{\Delta}{A}}{\ms{L}}\), \(\Gamma\)
splits into \(\Delta\) and \(\Xi\), and \(a\) is a pure term which typechecks in
\(\Xi\), then \(\lbrb{\ell}{a}\) targets \(\ms{L}\) given \(\Gamma\) is live on
entry. \rle{ite} states that, if \(\Gamma\) splits into \(\Delta\) and \(\Xi\),
\(e\) typechecks as a pure, boolean term in \(\Xi\), and that, given all
variables in \(\Delta\) are live on entry, \(s, t\) both target \(\ms{L}\), then
the if-statement \(\ite{e}{s}{t}\) targets \(\ms{L}\) given \(\Gamma\) is live
on entry. \rle{let-blk} and \rle{let2-blk} are quite similar to their term
analogues, except that the bound term \(e\) is allowed to be impure (otherwise,
programs could only make use of pure functions!). Finally, \rle{where} says
that, given a block \(t\) which targets \(\ms{L}\), and a CFG \(L\) with inputs
\(\ms{L}\) and outputs \(\ms{K}\), \(\ewhere{t}{L}\) targets \(\ms{K}\). CFGs
are typed using the typing rules \rle{nil-br}, which says that an empty
control-flow graph simply forwards the input labels directly to the output
labels, and \rle{cons-br}, which says that if \(L\) maps \(\ms{L}\) to either
\(\ms{K}\) or \(\lbl{\ell}[\Gamma](A)\), and \(t\) targets \(\ms{L}\) given that
\(\Gamma, \thyp{x}{A}{}\) is live on entry, then \(L\) maps \(\ms{L}\) to
\(\ms{K}\), having (potentially recursively) defined \(\lbl{\ell}\).

@@

To solve this problem, we need to introduce the concept of a
\textit{label-substitution} \(\mc{L}\), which we define as a function that takes
labels \(\lbl{\ell}\) and variables \(x\) to blocks \(t\) such that, for any
fresh variable \(x\), \([a/x]\mc{L}(\lbl{\ell}, x) = \mc{L}(\lbl{\ell}, a)\). We
introduce the judgement \(\lbsubst{\mc{L}}{\ms{L}}{\ms{K}}\) to state that the
label-substitution \(\mc{L}\) takes the label-set \(\ms{L}\) to the label-set
\(\ms{K}\); that is, for each \(\lbl{\ell}[\Gamma](x: A) \in \ms{L}\), we have
that \(\mc{L}(\lbl{\ell}, x)\) is a block that targets \(\ms{K}\). This is
formally defined via the rules \rle{lb-subst-nil}, which states that this is
vacuously true for \(\ms{L} = \cdot\), and \rle{lb-subst-cons}, which says that
if \(\lbsubst{\mc{L}}{\ms{L}}{\ms{K}}\) and \(\haslb{\Gamma,
\thyp{x}{A}{}}{\mc{L}(\lbl{\ell}, x)}{\ms{K}}\) then \(\lbsubst{\mc{L}}{\ms{L},
\lhyp{\lbl{\ell}}{\Gamma}{A}}{\ms{K}}\), in Figure \ref{fig:ssa-subst-typing}.
Given a context \(\gamma\) and a label-set \(\ms{L}\), we can convert it to a
label-substitution \(\substlbs{\gamma}{\ms{L}}\) using the definition in Figure
\ref{fig:ssa-subst-ops}: using this, we can state the full substitution lemma as
follows:
\begin{lemma}[Substitution] 
  Given \(\issubst{\gamma}{\Theta}{\Gamma}\),
  \begin{itemize}
    \item If \(\hasty{\Gamma}{p}{a}{A}{q}\), then
    \(\hasty{\Theta}{p}{[\gamma]a}{A}{q}\) and
    \(\hasty{\Theta}{p}{\exprletsubst{\substctx{\gamma}{\Gamma}}{a}}{A}{q}\)
    \item If \(\haslb{\Gamma}{t}{\ms{L}}\), then there exists \(\ms{K}\) such
    that \(\lbsubst{\substlbs{\gamma}{\ms{L}}}{\ms{K}}{\ms{L}}\) and
    \(\haslb{\Theta}{[\gamma]t}{\ms{K}}\),
    \(\haslb{\Theta}{\stmtletsubst{\substctx{\gamma}{\Gamma}}{t}}{\ms{K}}\)
    \item If \(\lhaslb{\ms{W}}{L}{\ms{L}}\), then there exists \(\ms{K},
    \ms{W}'\) such that \(\lbsubst{\substlbs{\gamma}{\ms{L}}}{\ms{K}}{\ms{L}}\),
    \(\lbsubst{\substlbs{\gamma}{\ms{W}'}}{\ms{W}}{\ms{W}'}\), and
    \(\lhaslb{\ms{W}'}{[\gamma]L}{\ms{K}}\)
  \end{itemize}
  Similarly, given \(\lbsubst{\mc{L}}{\ms{L}}{\ms{K}}\), we have that for all
  \(\haslb{\Gamma}{t}{\ms{L}}\), \(\haslb{\Gamma}{[\mc{L}]t}{\ms{K}}\)
\end{lemma}
